---
title: "Phaeo_Lightbox_Exp"
author: "Chris Carnivale"
date: "1/12/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(performance)
library(see)
library(ggplot2)
```

## Andrew Lightbox experiment

4 different light cycles to emulate different seasons of the polar regions



```{r reading in data, include=FALSE}
lightbox_phaeo <- read.csv("/Users/christophercarnivale/Desktop/Dissertation_data/Andrew's_data/LightboxPantarctica_2.csv")

LB_phaeo_colony <- read.csv("/Users/christophercarnivale/Desktop/Dissertation_data/Andrew's_data/LBData_Plus_Colony_2.csv")

#some quick making factors

lightbox_phaeo$treatment_time <- factor(lightbox_phaeo$treatment_time, levels = c("Initial", "14"))
LB_phaeo_colony$treatment_time <- factor(LB_phaeo_colony$treatment_time, levels = c("Initial", "14"))
```

```{r bacterivory tidying/calc w/o colony, include=FALSE}
lightbox_phaeo_ing <- lightbox_phaeo %>% 
  mutate(bac_conc_perML = bac_avg/bac_micro_filtered*dil_factor*cor_factor_grid,
         micro_conc_perML = micro_avg/bac_micro_filtered*dil_factor*cor_factor_grid,
         cells_conc_perML = cell_avg/count_filtered*dil_factor*cor_factor_grid,
         BM_ratio = bac_conc_perML/micro_conc_perML,
         BSP_ingested = (conc_bkgrd_per*BM_ratio) + conc_bkgrd_per,
         ing_percell_perhour = BSP_ingested*2)
lightbox_phaeo_ing_bygroup <- lightbox_phaeo_ing %>%
  group_by(treatment_time,treatment_light, treatment_nut) %>% 
  summarise(avg_ing = mean(ing_percell_perhour), 
            STD_ing = sd(ing_percell_perhour),
            STDerr_ing = STD_ing/sqrt(3),
            avg_cell_perML = mean(cells_conc_perML))
```

I need to double check and make sure that the bacteria slides were actually 2 ml
filtered. That seems REALLY high to me without some kind of dilution. However,
the trends should still shine through in the visualizations.

```{r bacterivory tidying/calc w colony, include=FALSE}
LB_phaeo_colony_ing <- LB_phaeo_colony %>% 
  mutate(bac_conc_perML = bac_avg/bac_micro_filtered*dil_factor*cor_factor_grid,
         micro_conc_perML = micro_avg/bac_micro_filtered*dil_factor*cor_factor_grid,
         cells_conc_perML = cell_avg/count_filtered*dil_factor*cor_factor_grid,
         BM_ratio = bac_conc_perML/micro_conc_perML,
         BSP_ingested = (conc_bkgrd_per*BM_ratio) + conc_bkgrd_per,
         ing_percell_perhour = BSP_ingested*2)

LB_phaeo_colony_ing_bygrp <- LB_phaeo_colony_ing %>%
  group_by(treatment_time,treatment_light, treatment_nut) %>% 
  summarise(avg_ing = mean(ing_percell_perhour), 
            STD_ing = sd(ing_percell_perhour),
            STDerr_ing = STD_ing/sqrt(3),
            avg_cell_perML = mean(cells_conc_perML))
```


```{r , include=FALSE}
ggplot(lightbox_phaeo_ing, aes(x = treatment_time, y = ing_percell_perhour))+
  geom_boxplot(aes(fill = treatment_nut))+
                 ylim(0,10)+
  facet_wrap(~treatment_light)

ggplot(lightbox_phaeo_ing, aes(x = treatment_time, y = cells_conc_perML))+
  geom_boxplot(aes(fill = treatment_nut))+
  facet_wrap(~treatment_light)

ggplot(LB_phaeo_colony_ing, aes(x = treatment_time, y = Colony_Area))+
  geom_boxplot(aes(fill = treatment_nut))+
  facet_wrap(~treatment_light)

ggplot(lightbox_phaeo_ing, aes(x = treatment_time, y = cells_conc_perML))+
  geom_point(aes(color = treatment_nut), position = position_dodge(0.5))+
  facet_wrap(~treatment_light)

ggplot(lightbox_phaeo_ing, aes(x = treatment_time, y = cells_conc_perML))+
  geom_point(aes(color = treatment_nut))+
  facet_wrap(~treatment_light)

ggplot(LB_phaeo_colony_ing, aes(x = treatment_time, y = Colony_Area))+
  geom_point(aes(color = treatment_nut))+
  facet_wrap(~treatment_light)
```

Need to make all plots boxplots (DONE)

Need to convert everything into total biomass - Average cell diameter is 6 um.
Average area per cell is 3^2pi or 9 pi um.

possibly calculate average growth rate or doubling time over the 2 week period?
just something to compare growth between light levels and nutrient with regards
to observed feeding behavior. MUCH better for statistical comparison than just 
raw biomass numbers.

```{r calculate biomass with conversion factor, include=FALSE}
LB_phaeo_colony_ing_totBIO <- mutate(LB_phaeo_colony_ing,
                                     Colony_cell_perML = Colony_Area/9/pi/2,
                                     Total_cell_perML = Colony_cell_perML+cells_conc_perML)

# 9 is radius squared and 2 is the volume of sample filtered in mL.

totBIO_box <- ggplot(LB_phaeo_colony_ing_totBIO, aes(x = treatment_time, y = Total_cell_perML))+
  geom_boxplot(aes(fill = treatment_nut))+
  facet_wrap(~treatment_light)

freeliving_box <- ggplot(lightbox_phaeo_ing, aes(x = treatment_time, y = ing_percell_perhour))+
  geom_boxplot(aes(fill = treatment_nut))+
                 ylim(0,10)+
  facet_wrap(~treatment_light)

totBIO_box
freeliving_box

quartile_table <- layer_data(totBIO_box)
#Need to add an ID row so he knows which quartiles correspond to what sample.
quartile_id_vector <- paste(LB_phaeo_colony_ing_totBIO$treatment_light, LB_phaeo_colony_ing_totBIO$treatment_time, LB_phaeo_colony_ing_totBIO$treatment_nut,sep = "_")
#This ID will work if I use a unique() function, however the order is incorrect. Might be easier to create a grouped dataset that corresponds to the grouping order used in geom_boxplot.
LB_phaeo_colony_ing_totBIO_grp <- group_by(LB_phaeo_colony_ing_totBIO,treatment_nut,treatment_light,treatment_time)

quartile_id_vector1 <- paste(LB_phaeo_colony_ing_totBIO_grp$treatment_light, LB_phaeo_colony_ing_totBIO_grp$treatment_time, LB_phaeo_colony_ing_totBIO_grp$treatment_nut,sep = "_")
unique(quartile_id_vector1)
#Reordering dataset to get the correct id order
LB_phaeo_colony_ing_totBIO_grp_reorder <- arrange(LB_phaeo_colony_ing_totBIO_grp, treatment_light)

quartile_id_vector2 <- paste(LB_phaeo_colony_ing_totBIO_grp_reorder$treatment_light, LB_phaeo_colony_ing_totBIO_grp_reorder$treatment_time, LB_phaeo_colony_ing_totBIO_grp_reorder$treatment_nut,sep = "_")

unique(quartile_id_vector2)

quartile_table$ID <- unique(quartile_id_vector2)

quartile_table[c(2:6,24)]
#lower = 1st quartile, middle = 2nd quartile, upper  = 3rd quartile, min and max are the top and bottom of the whiskers

quartile_table_ing <- layer_data(freeliving_box)

quartile_table_ing$ID <- unique(quartile_id_vector2)

quartile_table_ing[c(2:6,25)]

quartile_table_freecells <- layer_data(ggplot(lightbox_phaeo_ing, aes(x = treatment_time, y = cells_conc_perML))+
  geom_boxplot(aes(fill = treatment_nut))+
  facet_wrap(~treatment_light))

quartile_table_freecells$ID <- unique(quartile_id_vector2)

quartile_table_freecells[c(2:6,25)]

quartile_table_colonyArea <- layer_data(ggplot(LB_phaeo_colony_ing, aes(x = treatment_time, y = Colony_Area))+
  geom_boxplot(aes(fill = treatment_nut))+
  facet_wrap(~treatment_light))

quartile_table_colonyArea$ID <- unique(quartile_id_vector2)

quartile_table_colonyArea[c(2:6,24)]
```

```{r ANOVAs for Andrews dissertation, include=FALSE}
LB_phaeo_colony_ing_totBIO_T14 <- filter(LB_phaeo_colony_ing_totBIO, treatment_time == "14")

#full_nutrient_aov <- aov(ing_percell_perhour ~ treatment_light, filter(LB_phaeo_colony_ing_totBIO, treatment_nut == "100N"))

#summary(full_nutrient_aov)

#no_NP_aov <- aov(ing_percell_perhour ~ treatment_light, filter(LB_phaeo_colony_ing_totBIO, treatment_nut == "NP-"))

#summary(no_NP_aov)

#2-way ANOVA
#full_TWOway_aov <- aov(ing_percell_perhour ~ treatment_light*treatment_nut, LB_phaeo_colony_ing_totBIO_T14)

#summary(full_TWOway_aov)

#TukeyHSD(full_TWOway_aov)

#2-way ANOVA without 25 or 50% (This reduces the number of comparisons)
no50_25_TWOway_aov <- aov(ing_percell_perhour ~ treatment_light*treatment_nut, filter(LB_phaeo_colony_ing_totBIO_T14, treatment_nut != "25" & treatment_nut != "50"))

summary(no50_25_TWOway_aov)

TukeyHSD(no50_25_TWOway_aov)

#Checking normality
shapiro.test(LB_phaeo_colony_ing_totBIO_T14$ing_percell_perhour)

```

As I suspected, there's too many comparisons for a TUKEYS test. Not enough power to distinguish with significance the means between groups, but the 2 way ANOVA shows signals that nutrients matter apparently.

```{r One-way anova series for Publication, include=TRUE}
LB_phaeo_colony_ing_totBIO_T14_no50_no_25 <- filter(LB_phaeo_colony_ing_totBIO_T14, treatment_nut != "25" & treatment_nut != "50")

Light_ONEway_aov <- aov(ing_percell_perhour ~ treatment_light, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25, treatment_nut == "100N"))

summary(Light_ONEway_aov)

TukeyHSD(Light_ONEway_aov)

#Series of Nutrient One way anovas
Nut_allLight_ONEway_aov <- aov(ing_percell_perhour ~ treatment_nut, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25, treatment_light == "24H"))

Nut_dark_ONEway_aov <- aov(ing_percell_perhour ~ treatment_nut, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25, treatment_light == "24Z"))

Nut_14L_10Z_ONEway_aov <- aov(ing_percell_perhour ~ treatment_nut, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25, treatment_light == "14L.10Z"))

Nut_15H_7Z_ONEway_aov <- aov(ing_percell_perhour ~ treatment_nut, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25, treatment_light == "15H.2L.7Z"))

#Summaries of each
summary(Nut_allLight_ONEway_aov)

summary(Nut_dark_ONEway_aov)

summary(Nut_14L_10Z_ONEway_aov)

summary(Nut_15H_7Z_ONEway_aov)

#testing model residuals for normality
Nut_15H_7Z_ONEway_aov_resid <- resid(Nut_15H_7Z_ONEway_aov)

qqnorm(Nut_15H_7Z_ONEway_aov_resid)

plot(density(Nut_15H_7Z_ONEway_aov_resid))

check_model(Nut_15H_7Z_ONEway_aov)

check_model(Nut_allLight_ONEway_aov)
check_model(Nut_dark_ONEway_aov)

#The initial model are mostly normal...however I'm gonna normalize the data and test to see if it changes anything just incase jd asks for it.

LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm <- LB_phaeo_colony_ing_totBIO_T14_no50_no_25

LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm$ing_percell_perhour <- (LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm$ing_percell_perhour-mean(LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm$ing_percell_perhour))/sd(LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm$ing_percell_perhour)

shapiro.test(LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm$ing_percell_perhour)

Light_ONEway_aov_norm <- aov(ing_percell_perhour ~ treatment_light, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm, treatment_nut == "100N"))

summary(Light_ONEway_aov_norm)

TukeyHSD(Light_ONEway_aov_norm)

#Series of Nutrient One way anovas
Nut_allLight_ONEway_aov_norm <- aov(ing_percell_perhour ~ treatment_nut, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm, treatment_light == "24H"))

Nut_dark_ONEway_aov_norm <- aov(ing_percell_perhour ~ treatment_nut, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm, treatment_light == "24Z"))

Nut_14L_10Z_ONEway_aov_norm <- aov(ing_percell_perhour ~ treatment_nut, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm, treatment_light == "14L.10Z"))

Nut_15H_7Z_ONEway_aov_norm <- aov(ing_percell_perhour ~ treatment_nut, filter(LB_phaeo_colony_ing_totBIO_T14_no50_no_25_norm, treatment_light == "15H.2L.7Z"))

#Summaries of each
summary(Nut_allLight_ONEway_aov_norm)

summary(Nut_dark_ONEway_aov_norm)

summary(Nut_14L_10Z_ONEway_aov_norm)

summary(Nut_15H_7Z_ONEway_aov_norm)
```



